---
output:
  html_document:
    keep_md: yes
---
# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

1. I will use dplyr for data manipulation and ggplot2 for plotting.
2. Variable date in activity data frme need to be changed to Date object. 
3. Then I created a second data frame without NA value.
```{r}
library("dplyr")
library("ggplot2")

```
```{r}
activity <- read.csv("activity/activity.csv", header = TRUE, sep=",")
activity$date <- as.Date(activity$date) # change date variable to Date class
activity_no_rm <- activity[complete.cases(activity),] # Create data set without NA
options(digits=2)
options(scipen=999) # removing exponential printing numbers
```

## What is mean total number of steps taken per day?

First, using dplyr's group_by I want to group activity per day and then summarise using function sum all steps per each day.
```{r}
activity_per_day <- group_by(activity_no_rm, date)
steps_per_day <- summarise(activity_per_day,sumsteps = sum(steps))
```
Next we can plot a histogram...
```{r}
hist(steps_per_day$sumsteps
     , main = "Histogram with breaks = 20"
     , xlab = "Total steps per day"
     , col = "blue"
     , breaks = 20)
```
  
... and calculate the mean and median of total number of steps taken per day.
```{r}
activity_no_rm.mean <- mean(steps_per_day$sumsteps)
activity_no_rm.median <- median(steps_per_day$sumsteps)
```
The mean of total number of steps taken per day is `r activity_no_rm.mean` and the median is `r activity_no_rm.median`

## What is the average daily activity pattern?

To make such time series plot we need to group data frame with interval variable and summarise mean of steps per interval.  
Next I will plot a time series using ggplot2

```{r}
intervals_group <- group_by(activity_no_rm, interval)
steps_mean_per_interval <- summarise(intervals_group, meansteps=mean(steps))

g <- ggplot(steps_mean_per_interval
            , aes(x = interval, y = meansteps))
g + geom_line() +
        labs(y = "Average steps") +
        labs(x = " interval") +
        labs(title = "Average steps per interval")
```

Finding which 5 minute interval contains a maximum number of steps
```{r}
max_interval<-steps_mean_per_interval[which(steps_mean_per_interval$meansteps==max(steps_mean_per_interval$meansteps)),]
```
```{r}
interval.interval <-max_interval$interval
meanaveragesteps <- max_interval$meansteps
```

Interval `r interval.interval` has a maximum average steps of `r meanaveragesteps`.

## Imputing missing values

This time I need to use original activity data frame for count NA value in any variable.
```{r echo=TRUE}
colSums(is.na(activity))

```
There're 2304 missing values only in steps variable

My startegy is very simple: change NA value with the mean number of steps for the interval, across all days

```{r}
# impute values are from steps_mean_per_interval
impute <- steps_mean_per_interval
# using dplyr's left_join I will create a new data frame activityNEW with combined activity and steps_mean_per_interval,
# then creating index of NA in activeNEW data frame 
activityNEW <- left_join(activity, steps_mean_per_interval)
indexNA <- is.na(activityNEW$steps)

# Next I changed any NA with mean number of steps for the same interval
activityNEW$steps[indexNA]=activityNEW$meansteps[indexNA]

# only checking if there'are missing values
summary(activityNEW)
```

It seems that there's no NA value and all NA is filled with counted mean of steps per interval.
```{r}
activity_per_day_new <- group_by(activityNEW, date)
steps_per_day_new <- summarise(activity_per_day_new,sumsteps = sum(steps))
```
Next we can plot a histogram...
```{r}
hist(steps_per_day_new$sumsteps
     , main = "Histogram with breaks = 20"
     , xlab = "Total steps per day"
     , col = "blue"
     , breaks = 20)
```
  
... and calculate the mean and median of total number of steps taken per day
```{r}
activity_NEW.mean <- mean(steps_per_day_new$sumsteps)
activity_NEW.median <- median(steps_per_day_new$sumsteps)
```
The mean of total number of steps taken per day is `r activity_NEW.mean` and the median is `r activity_NEW.median`.  
Comparing to previously counted data without NA which mean of total number of steps taken per day was `r activity_no_rm.mean` and the median was `r activity_no_rm.median` we see that:  
1. The meanas have the same value  
2. The medians are different. For data with imputed NA the median is a little bit bigger.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
# For English weekdays
Sys.setlocale("LC_TIME", "English")
```
Next I create a factor variable with value "weekend" for "Saturday" and "Sunday" and "weekday" for anything else

```{r}
activityNEW$DayOrWeekend <- as.factor( 
               ifelse(
                       (weekdays(activityNEW$date)
                       %in% c("Saturday","Sunday"))
               ,"weekend","weekday"
               )
)

# just checking if factor variable was creating
table(activityNEW$DayOrWeekend)
```

```{r}
# First I group activityNEW by interval and DayOrWeekend factor 
# for summarise meansteps of interval for all weekday and weekend

intervals_DW_group <- group_by(activityNEW, interval, DayOrWeekend)
steps_mean_per_intervalDW <- summarise(intervals_DW_group, meansteps=mean(steps))

# Plot the graph 
ggplot(steps_mean_per_intervalDW, aes(x = interval, y= meansteps)) + geom_line() +facet_grid(DayOrWeekend ~.)
```


## End
